\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed

### 2014-10-08 Paper/grass.tex
All changes made
- Waiting for the green to change to another color
- Waiting for the last sentence in paragraph two.

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
\newcommand{\ubl}{U_\text{bl}}
\definecolor{tableShade}{gray}{0.8}

@@

This model has been applied successfully to ...!

@@

(ii)  The role of drag in damping the perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 

@@

@@

Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions from the solution of eq.~\eqref{Orr-somerfield}. 

@@

% We use a mean field model for the coupling between the flow and the vegetation. 
The vegetation drag on the flow is approximated by a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  where $\mathbf{f_{d}}$ is the drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit horizontal area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
This drag enters the governing equations of continuity and momentum balance, written as 
\begin{equation}
\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\label{navier-stokes}
\end{equation}
where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
$p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f_{d}}=C_N \rho u^{2}d\bx$, where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient. 
% We model it as 
% $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic). 
In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.
\newline
\begin{figure*}
 \includegraphics{SetAll_imgsc}
%\end{figure}
%\begin{figure*}
%\begin{tabular}{cccc}
%{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
%{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
%{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
%{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\

@@

The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, whereas the latter is due to the free interface. 
This expectations is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.

@@

Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.

@@

the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation 

@@

The growth rate $\sigma$ for a given wave number $k$ is the eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.

A threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition 
either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc}
assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.

Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of 
vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. The experimentally observed \monami ~wavelengths are not available for comparison.

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Dense grass limit &  & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~yes, near $y=\hg$~~			& ~~no, spans water column~~
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.

@@

Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
Because Mode 1 eigenfunction is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column. 
Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
The appearance of vegetation drag parameter in the dominant balances represented by eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
\newline
Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami} \cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tend to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case. 
\newline
The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
The rigidity of the vegetation can be evaluated using a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity. 
Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the buoyancy time scale to be about 2 s. 
Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
% Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 
We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.

In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
The eddy viscosity is also known to vary between the flow of unvegetated region and flow through the meadow \cite{Ghisal02}. 
Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
\newline
In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
We show the origin of this threshold using a consistent continuum hydrodynamic stability analysis for the shear flow that develops as a result of the vegetation-induced drag. 
Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
% The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow. 
Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

>>>>

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
\newcommand{\ubl}{U_\text{bl}}
\definecolor{tableShade}{gray}{0.8}

@@

This model has been applied successfully to predict the the observed frequency of \monami ...!

@@

(ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 

@@

We show that this drag results in an instability mode different from the Kelvin-Helmholtz instability.
\begin{figure}[b]

@@

Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p$ from the solution of eq.~\eqref{Orr-somerfield}. 

@@

The vegetation, assumed dense, exerts a drag modeled by a continuous body force $\mathbf{f}$.
% We assume the vegetation to be dense and model the drag by a continuous body force $\mathbf{f}$.
This drag enters the fluid mass and momentum balance equations as 
\begin{equation}
\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\label{navier-stokes}
\end{equation}
where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
$p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u^{2}d\bx$, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow. 
% We model it as 
% $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic). 
In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N \approx 1$.
\newline
\begin{figure*}
 \includegraphics{SetAll_imgsc}
%\end{figure}
%\begin{figure*}
%\begin{tabular}{cccc}
%{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
%{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
%{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
%{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\

@@

The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter is due to the free interface. 
This expectation is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.

@@

Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.

@@

the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation \cite{Drazin81} 

@@

The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.

A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
solution of eq.~\eqref{Orr-somerfield}. The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
The threshold Reynolds number increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc} assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.

Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
The experimentally observed \monami ~wavelengths are not available for comparison.

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Dense grass limit &  & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~yes, near $y=\hg$~~			& ~~no, spans water column~~
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.

@@

Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
Because Mode 1 eigenfunction is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column. 
Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
The appearance of vegetation drag parameter in the dominant balances represented by eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
\newline
Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami} \cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tend to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades. We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case. 
\newline
The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
The rigidity of the vegetation can be evaluated using a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity. 
Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the buoyancy time scale to be about 2 s. 
Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.

In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
Although these improvements in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
\newline
In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
We show the origin of this threshold using a consistent continuum hydrodynamic stability analysis for the shear flow that develops as a result of the vegetation-induced drag. 
Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
